---
title: "The 'gfilmm' package"
output: 
  rmarkdown::html_vignette:
    css: vignette.css
vignette: >
  %\VignetteIndexEntry{The 'gfilmm' package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  fig.width = 5, 
  fig.height = 4.5
)
```

## What does it do?

The 'gfilmm' package allows to generate simulations from the generalized 
fiducial distribution of the parameters of a Gaussian linear mixed model with 
categorical random effects (numeric random effects are not supported) and 
interval data. It also provides some helper functions to get summary statistics 
and confidence intervals.

The algorithm implemented in 'gfilmm' is the one described in the paper 
*Generalized fiducial inference for normal linear mixed models* written by 
Jessi Cisewski and Jan Hannig. It is coded in C++ and the code is based on the 
original Matlab code written by Jessi Cisewski.

Fiducial inference has something similar to Bayesian inference: the 
uncertainty about the parameters are represented by a distribution, the 
fiducial distribution, with the help of which we conduct inference on the 
parameters in a way similar to the Bayesian way, based on the posterior 
distribution of the parameters. The main difference is that there is no prior 
distribution (so fiducial inference is similar to *objective* Bayesian 
inference). The fiducial inference yields results close to the ones of the 
frequentist inference. 


## First example: a non-mixed linear model

The data must be given as a dataframe. Here we simulate data from a simple 
linear regression model:

```{r simulations_linreg}
set.seed(666L)
n <- 30L
x <- 1L:n
y <- rnorm(n, mean = x, sd = 2)
y_rounded <- round(y, digits = 1L)
dat <- data.frame(
  ylwr = y_rounded - 0.05,
  yupr = y_rounded + 0.05,
  x = x
)
```

Now we run the fiducial sampler:

```{r gfi_linreg}
library(gfilmm)
fidSims <- gfilmm(
  y = ~ cbind(ylwr, yupr), # interval data
  fixed = ~ x,             # fixed effects
  random = NULL,           # random effects
  data = dat,              # data
  N = 20000L               # number of simulations
)
```

A summary of the fiducial simulations (the `Pr(=0)` column will be explained 
latter):

```{r gfiSummary_linreg}
gfiSummary(fidSims)
```

The fiducial confidence intervals are close to the frequentist ones:

```{r lm_linreg}
lmfit <- lm(y ~ x)
confint(lmfit)
```

The fiducial cumulative distribution function of the slope:

```{r gfiCDF_linreg}
Fslope <- gfiCDF(~ x, fidSims)
plot(Fslope, main = "Slope", ylab = expression("Pr("<="x)"))
```

To get a fiducial density, I recommend the 'kde1d' package:

```{r gfiDensity_linreg}
library(kde1d)
kfit <- kde1d(fidSims$VERTEX[["x"]], weights = fidSims$WEIGHT, mult = 4)
curve(dkde1d(x, kfit), from = 0.7, to = 1.1)
```

***Fiducial predictive inference***. The `gfilmmPredictive` function samples 
the generalized fiducial predictive distribution. All the functions seen above 
can be applied to the output.

```{r gfilmmPredictive}
fpd <- gfilmmPredictive(fidSims, newdata = data.frame(x = c(1, 30)))
gfiSummary(fpd)
```

Compare with the frequentist approach:

```{r lmpredict}
predict(lmfit, newdata = data.frame(x = c(1, 30)), interval = "prediction")
```



## A mixed model

Now let us simulate some data from a one-way ANOVA model with a random factor:

```{r simulations_AOV1R}
mu           <- 10000 # grand mean
sigmaBetween <- 2
sigmaWithin  <- 3
I            <- 6L # number of groups
J            <- 5L # sample size per group

set.seed(31415926L)
groupmeans <- rnorm(I, mu, sigmaBetween)
y          <- c(
  vapply(groupmeans, function(gmean) rnorm(J, gmean, sigmaWithin), numeric(J))
)
y_rounded  <- round(y, digits = 1L)
dat        <- data.frame(
                ylwr = y_rounded - 0.05,
                yupr = y_rounded + 0.05,
                group = gl(J, I)
              )
```

We run the fiducial sampler:

```{r gfilmm_AOV1R}
fidSims <- gfilmm(~ cbind(ylwr, yupr), ~ 1, ~ group, data = dat, N = 20000L)
```

Observe that the between standard deviation `sigma_group` has a positive value 
in the `Pr(=0)` column:

```{r gfiSummary_AOV1R}
gfiSummary(fidSims)
```

What does it mean? The fiducial distributions of the variance components have a 
mass at zero, and this value is the probability that the between standard 
deviation equals zero. So you have to be careful if you are interested in the 
fiducial density of a standard deviation: if `Pr(=0)` is not null for the 
standard deviation you are interested in, the fiducial distribution of this 
standard deviation *does not have a density*. It has a mass at zero, and a 
density on the strictly positive real numbers.

Compare the fiducial confidence interval of the grand mean to its Kenward-Roger 
confidence interval:

```{r KenwardRoger, message=FALSE}
library(lmerTest)
library(emmeans)
fit <- lmer(y ~ (1|group), data = dat)
emmeans(fit, ~ 1)
```

Actually the design is balanced, and we can get the same confidence interval by 
the frequentist method:

```{r AOV1R}
library(AOV1R)
aovfit <- aov1r(y ~ dat$group)
confint(aovfit)
```

With `gfiConfInt` we can get a fiducial confidence interval for *any* parameter 
of interest, for example the coefficient of total variation:

```{r gfiConfInt_CV_AOV1R}
gfiConfInt(~ sqrt(sigma_group^2 + sigma_error^2)/`(Intercept)`, fidSims)
```


