---
title: "The 'gfilmm' package"
output: 
  rmarkdown::html_vignette:
    css: vignette.css
vignette: >
  %\VignetteIndexEntry{The 'gfilmm' package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# What does it do?

The 'gfilmm' package allows to generate simulations from the generalized 
fiducial distribution of the parameters of a Gaussian linear mixed model with 
categorical random effects (numeric random effects are not supported) and 
interval data. It also provides some helper functions to get summary statistics 
and confidence intervals.

The algorithm implemented in 'gfilmm' is the one described in the paper 
*Generalized fiducial inference for normal linear mixed models* written by 
Jessi Cisewski and Jan Hannig. It is coded in C++ and the code is based on the 
original Matlab code written by Jessi Cisewski.

# First example: a non-mixed linear model

The data must be given as a dataframe. Here we simulate data from a simple 
linear regression model:

```{r simulations_linreg}
set.seed(666)
n <- 30L
x <- 1L:n
y <- rnorm(n, mean = x, sd = 2)
y_rounded <- round(y, digits = 1L)
dat <- data.frame(
  ylwr = y_rounded - 0.05,
  yupr = y_rounded + 0.05,
  x = x
)
```

Now we run the fiducial sampler:

```{r gfi_linreg}
library(gfilmm)
fidSims <- gfilmm(
  y = ~ cbind(ylwr, yupr), # interval data
  fixed = ~ x,             # fixed effects
  random = NULL,           # random effects
  data = dat,              # data
  N = 500                # number of simulations
)
```

A summary of the fiducial simulations (the `Pr(=0)` column will be explained 
latter):

```{r gfiSummary_linreg}
gfiSummary(fidSims)
```

The fiducial confidence intervals are close to the frequentist ones:

```{r lm_linreg}
confint(lm(y ~ x))
```

The fiducial cumulative distribution function of the slope:

```{r gfiCDF_linreg}
Fslope <- gfiCDF(~ x, fidSims)
plot(Fslope, main = "Slope", ylab = expression("Pr("<="x)"))
```

To get a fiducial density, I recommend the 'kde1d' package:

```{r gfiDensity_linreg}
library(kde1d)
kfit <- kde1d(fidSims$VERTEX[["x"]], weights = fidSims$WEIGHT, mult = 4)
curve(dkde1d(x, kfit), from = 0.7, to = 1.1)
```

# A mixed model

Now let us simulate some data from a one-way ANOVA model with a random factor:

```{r simulations_AOV1R}
mu           <- 10000 # grand mean
sigmaBetween <- 2
sigmaWithin  <- 3
sigmaTotal   <- sqrt(sigmaBetween^2 + sigmaWithin^2)
CV           <- sigmaTotal / mu # coefficient of total variation
n            <- 8L # sample size per group

set.seed(666)
groupmeans <- rnorm(2L, mu, sigmaBetween)
y1         <- rnorm(n, groupmeans[1L], sigmaWithin) 
y2         <- rnorm(n, groupmeans[2L], sigmaWithin) 
y          <- c(y1, y2)
y_rounded  <- round(c(y1, y2), digits = 1L)
dat        <- data.frame(
                ylwr = y_rounded - 0.05,
                yupr = y_rounded + 0.05,
                group = gl(2L, n)
              )
```

We run the fiducial sampler:

```{r gfilmm_AOV1R}
fidSims <- gfilmm(~ cbind(ylwr, yupr), ~ 1, ~ group, data = dat, N = 500L)
```

Observe that the between standard deviation `sigma_group` has a positive value 
in the `Pr(=0)` column:

```{r gfiSummary_AOV1R}
gfiSummary(fidSims)
```

What does it mean? The fiducial distributions of the variance components have a 
mass at zero, and this value is the probability that the between standard 
deviation equals zero.

Compare the fiducial confidence interval of the grand mean to its Kenward-Roger 
confidence interval:

```{r KenwardRoger, message=FALSE}
library(lmerTest)
library(emmeans)
fit <- lmer(y ~ (1|group), data = dat)
emmeans(fit, ~ 1)
```

With `gfiConfInt` we can get a fiducial confidence interval for *any* parameter 
of interest, for example the coefficient of total variation:

```{r gfiConfInt_CV_AOV1R}
gfiConfInt(~ sqrt(sigma_group^2 + sigma_error^2)/`(Intercept)`, fidSims)
```

